### 2023-02-17 chapters/interpretability/relaxed-cnf/experiment.tex
formatting
highlevel pass
connecting chapters

%%\vspace{-3pt}
\subsection{Experiment Methodology}
We perform experiments on a high-performance computer cluster, where each node consists of E$ 5$-$2690\text{ v}3 $ CPU with $ 24 $ cores, $ 96 $ GB of RAM. Each experiment is run on four cores of a node with $ 16 $ GB memory.   We compare the performance of  $ \crr $   with  state-of-the-art classifiers, e.g. $ \IMLI  $~\cite{GM2019}, RIPPER~\cite{C1995}, BRS~\cite{WRDLKM2017},  random forest (RF), support vector classifier (SVC), nearest neighbors classifiers ($ k $-NN), and $ l_1 $ penalized logistic regression (LR). Among them, {\IMLI}, BRS, and RIPPER are rule-based classifiers. In particular,
{\IMLI} generates classification rules in CNF using a MaxSAT-based formulation and we use Open-WBO \cite{MML2014} as the MaxSAT solver for {\IMLI}. We compare with propositional rule learning algorithm RIPPER, which is implemented in WEKA~\cite{HFHPRW2009} and generates classification rules in the form of decision lists. BRS is a Bayesian framework for generating rule sets expressed as DNF. For  other classifiers, we use the Scikit-learn module of Python~\cite{PVGMTGB2011}. For all  classifiers, we set the training cut-off time to $ 1500 $  {seconds}.

\begin{table}
	\caption[Accuracy, rule-size, and training time of rule-based classifiers]{Comparisons of test accuracy, rule-size and training time among different rule-based classifiers. Every cell in the last four columns contains the test accuracy in percentage (top value), rule size  (middle value), and  training time in seconds (bottom value). In the experiments, {\crr} shows higher test accuracy than IMLI and generates sparser rules than RIPPER. Number in bold denotes the best result, such as maximum test accuracy, minimum rule-size, and minimum training time among competitive classifiers.}
	\label{interpretability_crr_tab:rule_based_classifiers}
%	\setlength{\tabcolsep}{3pt} 
	\begin{center}
		\begin{tabular}{l  r  r r r r r r r rrr}
			\toprule
			{{Dataset}} & Size & Features  & RIPPER & BRS & IMLI & CRR\\
			\midrule
			{ Heart}   & $  303 $  & $  31 $  & $   \textbf{78.69}  $    & $   72.13  $    & $   72.13  $    & $   77.69  $   \\ & & 
			& $   7 $    & $   19 $    & $   13 $    & $   \textbf{4.5}  $   \\ & & 
			& $   5.27 \text{s}  $    & $   25.07 \text{s}  $    & $   \textbf{1.8} \text{s}  $    & $   122.5 \text{s}  $   \\\\
			{ Ionosphere}   & $  351 $  & $  144 $  & $   88.65  $    & $   \textbf{91.43}  $    & $   89.29  $    & $   \textbf{91.43}  $   \\ & & 
			& $   8 $    & $   \textbf{4} $    & $   8.5  $    & $   20 $   \\ & & 
			& $   5.87 \text{s}  $    & $   75.53 \text{s}  $    & $   \textbf{2.09} \text{s}  $    & $   5.59 \text{s}  $   \\\\ 
			{ WDBC}   & $  569 $  & $  88 $  & $   95.22  $    & $   \textbf{95.65}  $    & $   93.91  $    & $   94.69  $   \\ & & 
			& $   7.5  $    & $   12 $    & $   \textbf{7} $    & $   34.5  $   \\ & & 
			& $   5.7 \text{s}  $    & $   630.23 \text{s}  $    & $   \textbf{1.38} \text{s}  $    & $   316.32 \text{s}  $   \\\\ 
			{ Magic}   & $  19020 $  & $  79 $  & $   \textbf{84.04}  $    & $   74.15  $    & $   71.97  $    & $   81.31  $   \\ & & 
			& $   115 $    & $   \textbf{3} $    & $   24 $    & $   31 $   \\ & & 
			& $   \textbf{15.86} \text{s}  $    & $   56.46 \text{s}  $    & $   141.2 \text{s}  $    & $   1012.6 \text{s}  $   \\\\ 
			{ Tom's HW}   & $  28179 $  & $  910 $  & $   \textbf{97.4}  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   95.88  $    & $   97.34  $   \\ & & 
			& $   36 $    &       & $   30 $    & $   \textbf{4} $   \\ & & 
			& $   \textbf{42.73} \text{s}  $    &       & $   92.65 \text{s}  $    & $   1071.58 \text{s}  $   \\\\ 
			{ Credit}   & $  30000 $  & $  110 $  & $   81.68  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   81.42  $    & $  \textbf{ 82.04}  $   \\ & & 
			& $   38.5  $    &       & $   \textbf{10} $    & $   32 $   \\ & & 
			& $   \textbf{14.52} \text{s}  $    &       & $   17.66 \text{s}  $    & $   1021.35 \text{s}  $   \\\\ 
			{ Adult}   & $  32561 $  & $  144 $  & $   84.31  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   82.08  $    & $   \textbf{84.86}  $   \\ & & 
			& $   94 $    &       & $   23 $    & $   \textbf{18} $   \\ & & 
			& $   27.61 \text{s}  $    &       & $   \textbf{11.91} \text{s}  $    & $   1016.36 \text{s}  $   \\\\ 
			{ Twitter}   & $  49999 $  & $  1511 $  & $   \textbf{95.74}  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   94.24  $    & $   95.16  $   \\ & & 
			& $   179.5  $    &       & $   57 $    & $   \textbf{12} $   \\ & & 
			& $   \textbf{170.87} \text{s}  $    &       & $   238.29 \text{s}  $    & $   1144.66 \text{s}  $   \\\\ 
			{ Weather-AUS}   & $  107696 $  & $  141 $  & $   \textbf{84.57}  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   82.83  $    & $   83.34  $   \\ & & 
			& $   195 $    &       & $   7 $    & $   \textbf{2} $   \\ & & 
			& $   \textbf{121.02} \text{s}  $    &       & $   366.12 \text{s}  $    & $   1115.27 \text{s}  $   \\\\ 
			{ Skin}   & $  245057 $  & $  119 $  & $   98.32  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   \textbf{98.92}  $    & $   95.08  $   \\ & & 
			& $   725 $    &       & $   201 $    & $   \textbf{29} $   \\ & & 
			& $   1313.19 \text{s}  $    &       & $   \textbf{103.8} \text{s}  $    & $   825.6 \text{s}  $   \\\\ 
			\bottomrule
		\end{tabular}
	\end{center}

@@

\begin{figure}[t]
	\centering

>>>>

%%\vspace{-3pt}
\subsection{Experiment Methodology}
We perform experiments on a high-performance computer cluster, where each node consists of E$ 5$-$2690\text{ v}3 $ CPU with $ 24 $ cores, $ 96 $ GB of RAM. Each experiment is run on four cores of a node with $ 16 $ GB memory.   We compare the performance of  $ \crr $   with  state-of-the-art classifiers, e.g. $ \IMLI  $~\cite{GM2019}, RIPPER~\cite{cohen1995fast}, BRS~\cite{wang2017bayesian},  random forest (RF), support vector classifier (SVC), nearest neighbors classifiers ($ k $-NN), and $ l_1 $ penalized logistic regression (LR). Among them, {\IMLI}, BRS, and RIPPER are rule-based classifiers. In particular,
{\IMLI} generates classification rules in CNF using a MaxSAT-based formulation and we use Open-WBO \cite{martins2014open} as the MaxSAT solver for {\IMLI}. We compare with propositional rule learning algorithm RIPPER, which is implemented in WEKA~\cite{HFHPRW2009} and generates classification rules in the form of decision lists. BRS is a Bayesian framework for generating rule sets expressed as DNF. For  other classifiers, we use the Scikit-learn module of Python~\cite{scikit-learn}. For all  classifiers, we set the training cut-off time to $ 1500 $  {seconds}.

\begin{table}
	\caption[Accuracy, rule-size, and training time of rule-based classifiers]{Comparisons of test accuracy, rule-size and training time among different rule-based classifiers. Every cell in the last four columns contains the test accuracy in percentage (top value), rule size  (middle value), and  training time in seconds (bottom value). In the experiments, {\crr} shows higher test accuracy than IMLI and generates sparser rules than RIPPER. Number in bold denotes the best result, such as maximum test accuracy, minimum rule-size, and minimum training time among competitive classifiers.}
	\label{interpretability_crr_tab:rule_based_classifiers}

%	\setlength{\tabcolsep}{3pt} 
	\begin{center}
		\begin{tabular}{l  r  r r r r r r r rrr}
			\toprule
			{{Dataset}} & Size & Features  & RIPPER & BRS & IMLI & CRR\\
			\midrule
			{ Heart}   & $  303 $  & $  31 $  & $   \textbf{78.69}  $    & $   72.13  $    & $   72.13  $    & $   77.69  $   \\ & & 
			& $   7 $    & $   19 $    & $   13 $    & $   \textbf{4.5}  $   \\ & & 
			& $   5.27 \text{s}  $    & $   25.07 \text{s}  $    & $   \textbf{1.8} \text{s}  $    & $   122.5 \text{s}  $   \\
			\midrule 
			{ Ionosphere}   & $  351 $  & $  144 $  & $   88.65  $    & $   \textbf{91.43}  $    & $   89.29  $    & $   \textbf{91.43}  $   \\ & & 
			& $   8 $    & $   \textbf{4} $    & $   8.5  $    & $   20 $   \\ & & 
			& $   5.87 \text{s}  $    & $   75.53 \text{s}  $    & $   \textbf{2.09} \text{s}  $    & $   5.59 \text{s}  $   \\
			\midrule 
			{ WDBC}   & $  569 $  & $  88 $  & $   95.22  $    & $   \textbf{95.65}  $    & $   93.91  $    & $   94.69  $   \\ & & 
			& $   7.5  $    & $   12 $    & $   \textbf{7} $    & $   34.5  $   \\ & & 
			& $   5.7 \text{s}  $    & $   630.23 \text{s}  $    & $   \textbf{1.38} \text{s}  $    & $   316.32 \text{s}  $   \\
			\midrule 
			{ Magic}   & $  19020 $  & $  79 $  & $   \textbf{84.04}  $    & $   74.15  $    & $   71.97  $    & $   81.31  $   \\ & & 
			& $   115 $    & $   \textbf{3} $    & $   24 $    & $   31 $   \\ & & 
			& $   \textbf{15.86} \text{s}  $    & $   56.46 \text{s}  $    & $   141.2 \text{s}  $    & $   1012.6 \text{s}  $   \\
			\midrule 
			{ Tom's HW}   & $  28179 $  & $  910 $  & $   \textbf{97.4}  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   95.88  $    & $   97.34  $   \\ & & 
			& $   36 $    &       & $   30 $    & $   \textbf{4} $   \\ & & 
			& $   \textbf{42.73} \text{s}  $    &       & $   92.65 \text{s}  $    & $   1071.58 \text{s}  $   \\
			\midrule 
			{ Credit}   & $  30000 $  & $  110 $  & $   81.68  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   81.42  $    & $  \textbf{ 82.04}  $   \\ & & 
			& $   38.5  $    &       & $   \textbf{10} $    & $   32 $   \\ & & 
			& $   \textbf{14.52} \text{s}  $    &       & $   17.66 \text{s}  $    & $   1021.35 \text{s}  $   \\
			\midrule 
			{ Adult}   & $  32561 $  & $  144 $  & $   84.31  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   82.08  $    & $   \textbf{84.86}  $   \\ & & 
			& $   94 $    &       & $   23 $    & $   \textbf{18} $   \\ & & 
			& $   27.61 \text{s}  $    &       & $   \textbf{11.91} \text{s}  $    & $   1016.36 \text{s}  $   \\
			\midrule 
			{ Twitter}   & $  49999 $  & $  1511 $  & $   \textbf{95.74}  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   94.24  $    & $   95.16  $   \\ & & 
			& $   179.5  $    &       & $   57 $    & $   \textbf{12} $   \\ & & 
			& $   \textbf{170.87} \text{s}  $    &       & $   238.29 \text{s}  $    & $   1144.66 \text{s}  $   \\
			\midrule 
			{ Weather-AUS}   & $  107696 $  & $  141 $  & $   \textbf{84.57}  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   82.83  $    & $   83.34  $   \\ & & 
			& $   195 $    &       & $   7 $    & $   \textbf{2} $   \\ & & 
			& $   \textbf{121.02} \text{s}  $    &       & $   366.12 \text{s}  $    & $   1115.27 \text{s}  $   \\
			\midrule 
			{ Skin}   & $  245057 $  & $  119 $  & $   98.32  $    &   \multicolumn{1}{c}{\multirow{3}{*}{\textemdash}}       & $   \textbf{98.92}  $    & $   95.08  $   \\ & & 
			& $   725 $    &       & $   201 $    & $   \textbf{29} $   \\ & & 
			& $   1313.19 \text{s}  $    &       & $   \textbf{103.8} \text{s}  $    & $   825.6 \text{s}  $   \\
			\bottomrule
		\end{tabular}
	\end{center}

@@

\begin{figure}
	\centering

### 2023-02-17 chapters/interpretability/relaxed-cnf/ilp_formulation.tex
with new formatting
connecting chapters

\section{{\crr}: Classification Rules in Relaxed Logical Form}
\label{interpretability_crr_sec:ilp}
In this section, we describe the main contribution of our work, {\crr}, a  framework for learning   relaxed-CNF rules.  {\crr}  converts the learning problem into  an ILP-based formulation,  learns the optimal assignment of variables and constructs rule $ \Rule $ based on the assignment. We organize the rest of this section as follows.
We  discuss the decision variables in Section ~\ref{interpretability_crr_sec:variables},  the constraints and the linear programming relaxation in Section~\ref{interpretability_crr_sec:ilp_query}, the incremental learning in Section~\ref{interpretability_crr_sec:inc_ilp}, feature discretization in Section~\ref{interpretability_crr_sec:discretization}, and adaptation of {\crr} for learning other classification rules in Section~\ref{interpretability_crr_sec:other_rules}.
% We conclude this section by describing how {\crr} can be adjusted to learn interpretable rules in other logical forms in Section~\ref{interpretability_crr_sec:other_format}.

>>>>

\section{Classification Rules in Relaxed Logical Form}
\label{interpretability_crr_sec:ilp}
In this section, we describe the main contribution of this chapter, {\crr}, a  framework for learning   relaxed-CNF rules.  {\crr}  converts the learning problem into  an ILP-based formulation,  learns the optimal assignment of variables and constructs rule $ \Rule $ based on the assignment. We organize the rest of this section as follows.
We  discuss the decision variables in Section ~\ref{interpretability_crr_sec:variables},  the constraints and the linear programming relaxation in Section~\ref{interpretability_crr_sec:ilp_query}, the incremental learning in Section~\ref{interpretability_crr_sec:inc_ilp}, feature discretization in Section~\ref{interpretability_crr_sec:discretization}, and adaptation of {\crr} for learning other classification rules in Section~\ref{interpretability_crr_sec:other_rules}.
% We conclude this section by describing how {\crr} can be adjusted to learn interpretable rules in other logical forms in Section~\ref{interpretability_crr_sec:other_format}.

### 2023-02-17 chapters/interpretability/relaxed-cnf/introduction.tex
formatting
connecting chapters

\label{chapter:crr}
The widespread adoption of prediction systems in various safety-critical domains such as medical diagnosis, law, education, and many others has led to the increased importance of presenting the decision functions in interpretable representations~\cite{K2001,MVBB2005,SAD2015,S2014,TV2013}.  To enable  safe, robust, and trustworthy integration of such systems, the end users require them to support interpretability, privacy, and fairness in decision-making~\cite{DF2018,VML2012,WRLKM2015,ZUR2017}. In this context,  rule-based representations are considered interpretable in presenting the decision functions to users~\cite{LKCL2019,WR2015,WRDLKM2017}. A recent body of work has studied sparsity-inducing objectives for classification rules in CNF or DNF and demonstrated that they often achieve high interpretability\textemdash defined in terms of rule-sparsity\textemdash with a minimal sacrifice in classification accuracy ~\cite{GM2019,LKCL2019,MM18}. Although CNF/DNF rules are considered interpretable, they are less expressive compared to Boolean cardinality constraints in propositional logic. A Boolean cardinality constraint allows one to express numerical bounds on Boolean variables~\cite{sinz2005towards}. In this work, we introduce a novel formulation of interpretable classification rules, namely \emph{relaxed-CNF}, which achieve benefits of  both worlds: it is interpretable similar to CNF/DNF  but more expressible by allowing cardinality constraints in the representation.

@@

We now provide an introduction to relaxed-CNF rules. In interpretable rule-based classification, the simplest logical rules are single level-rules: ORs or ANDs of Boolean \textit{literals}, where each literal denotes either a Boolean input feature or its negation. A \emph{clause} is a collection of $ N $ literals connected by  OR/AND. To satisfy a clause, an OR operator requires $ 1 $ out of $ N $ literals to be assigned to  $ 1 $ in the clause, while an AND operator requires all $N$ out of $N$ literals to be assigned to  $ 1 $. A CNF   formula  is a conjunction (AND) of {clauses} where each clause is a disjunction (OR) of {literals}, and  a DNF  formula is a disjunction of clauses where each clause is a conjunction of literals. Therefore, CNF  and DNF  formulas can be viewed as two-level rules with several applications in interpretable decision-making. For example, a decision set is a DNF rule referring to a set of ``if-else'' conditions~\cite{ignatiev2018sat,lakkaraju2016interpretable}.  In this chapter, we consider a richer set of logical formulas that capture the structure of checklists. To this end, we consider \emph{hard}-OR clauses, where at least $ M > 1$ out of $ N $ literals are assigned to $ 1 $, and we similarly define {\emph{soft}-AND} clauses which allow some of the literals (at most $ N-M $) to be $ 0 $. To be precise,  the definitions of hard-OR and soft-AND overlap. Consequently, we use hard-OR when $M \le N/2$ and soft-AND otherwise. To extend the standard definition of CNF (which is ANDs of ORs),  we define \emph{relaxed}-CNF to denote soft-ANDs of hard-ORs. Similarly, \emph{relaxed}-DNF is hard-ORs of soft-ANDs. Since   hard-OR and soft-AND are differentiated based on the value of $ M $ and $ N $, relaxed-CNF and relaxed-DNF have the same structural representation. In early work, Craven and Shavlik~\cite{craven1996extracting} considered single level $ M $-of-$ N $ rules to explain black-box neural-network classifiers. Recently, Emad et al.~\cite{EVM2015} have developed a semi-quantitative group testing approach for learning sparse single level $M$-of-$N$ rules, which are quite restrictive in their ability to fit the data. In contrast, in this work, we study a much richer family of two-level relaxed-CNF rules.

\begin{figure*}
	\begin{center}
		\includegraphics[scale=.65]{figures/interpretability/relaxed-cnf/checklist.pdf}
	\end{center}
	\caption[Illustration of a relaxed-CNF classification rule]{An illustrative example of a  relaxed CNF classification rule, which describes the decision function in the form of a  two-level checklist. This classifier is learned on the WDBC (Wisconsin diagnostic breast cancer) dataset and it predicts whether a tumor cell is malignant or not based on the characteristics of the  tumor cell. The first column in each checklist contains Boolean literals. An entry in the second column is  $ 1 $ if the corresponding literal is true by an observed tumor cell, and $ 0 $ otherwise. In the first level, checklist A (resp.\ B) is true if the count of true  literals is at least $ 2 $ (resp.\ $ 3 $). In the second level, a tumor cell is predicted as malignant if the count of true checklists is at least $ 1 $. }
	\label{interpretability_crr_exmpl:wdbc}
\end{figure*}

@@

{Through a comprehensive experimental evaluation over datasets from the UCI and Kaggle repository, we observe that {\crr} with relaxed-CNF rules achieves an improved trade-off between accuracy and rule sparsity and scales to datasets with more than $ 10^5 $ samples. More significantly, {\crr} generates relaxed-CNF rules with higher accuracy than CNF rules generated by~\cite{GM2019}. Furthermore, compared to decision lists generated by~\cite{C1995}, relaxed-CNF rules are sparser in large datasets.}
%can also reach the same level of accuracy compared to plain CNF/DNF rules with smaller relaxed-CNF rules. 
%
%We have conducted a comprehensive experimental study on the datasets from the UCI repository. Our empirical study demonstrates  an increase in the prediction accuracy for choosing relaxed-CNF rules over regular CNF rules. The incremental approach achieves runtime scalability without loss of accuracy.

>>>>

\label{chapter:crr}
%The widespread adoption of prediction systems in various safety-critical domains such as medical diagnosis, law, education, and many others has led to the increased importance of presenting the decision functions in interpretable representations~\cite{K2001,MVBB2005,SAD2015,S2014,TV2013}.  To enable  safe, robust, and trustworthy integration of such systems, the end users require them to support interpretability, privacy, and fairness in decision-making~\cite{DF2018,VML2012,WRLKM2015,ZUR2017}. In this context,  rule-based representations are considered interpretable in presenting the decision functions to users~\cite{LKCL2019,WR2015,WRDLKM2017}. A recent body of work has studied sparsity-inducing objectives for classification rules in CNF or DNF and demonstrated that they often achieve high interpretability\textemdash defined in terms of rule-sparsity\textemdash with a minimal sacrifice in classification accuracy ~\cite{GM2019,LKCL2019,MM18}.

In this chapter, we focus on improving the expressiveness of interpretable rule-based classifiers. Continuing from chapter~\ref{chapter:imli},
although CNF/DNF rules are considered interpretable, they are less expressive compared to Boolean cardinality constraints in propositional logic. A Boolean cardinality constraint allows one to express numerical bounds on Boolean variables~\cite{sinz2005towards}. In this chapter, we introduce a novel formulation of interpretable classification rules, namely \emph{relaxed-CNF}, which achieve benefits of  both worlds: it is interpretable similar to CNF/DNF  but more expressible by allowing cardinality constraints in the representation.

@@

We now provide an introduction to relaxed-CNF rules. In interpretable rule-based classification, the simplest logical rules are single level-rules: ORs or ANDs of Boolean \textit{literals}, where each literal denotes either a Boolean input feature or its negation. A \emph{clause} is a collection of $ N $ literals connected by  OR/AND. To satisfy a clause, an OR operator requires $ 1 $ out of $ N $ literals to be assigned to  $ 1 $ in the clause, while an AND operator requires all $N$ out of $N$ literals to be assigned to  $ 1 $. A CNF   formula  is a conjunction (AND) of {clauses} where each clause is a disjunction (OR) of {literals}, and  a DNF  formula is a disjunction of clauses where each clause is a conjunction of literals. Therefore, CNF  and DNF  formulas can be viewed as two-level rules with several applications in interpretable decision-making. For example, a decision set is a DNF rule referring to a set of ``if-else'' conditions~\cite{ignatiev2018sat,lakkaraju2016interpretable}.  In this chapter, we consider a richer set of logical formulas that capture the structure of checklists. To this end, we consider \emph{hard}-OR clauses, where at least $ M > 1$ out of $ N $ literals are assigned to $ 1 $, and we similarly define {\emph{soft}-AND} clauses which allow some of the literals (at most $ N-M $) to be $ 0 $. To be precise,  the definitions of hard-OR and soft-AND overlap. Consequently, we use hard-OR when $M \le N/2$ and soft-AND otherwise. To extend the standard definition of CNF (which is ANDs of ORs),  we define \emph{relaxed}-CNF to denote soft-ANDs of hard-ORs. Similarly, \emph{relaxed}-DNF is hard-ORs of soft-ANDs. Since   hard-OR and soft-AND are differentiated based on the value of $ M $ and $ N $, relaxed-CNF and relaxed-DNF have the same structural representation. In early work, Craven and Shavlik~\cite{craven1996extracting} considered single level $ M $-of-$ N $ rules to explain black-box neural-network classifiers. Recently, Emad et al.~\cite{EVM2015} have developed a semi-quantitative group testing approach for learning sparse single level $M$-of-$N$ rules, which are quite restrictive in their ability to fit the data. In contrast, in this chapter, we study a much richer family of two-level relaxed-CNF rules.

\begin{figure*}
	\begin{center}
		\includegraphics[scale=.62]{figures/interpretability/relaxed-cnf/checklist.pdf}
	\end{center}
	\caption[Illustration of a relaxed-CNF classification rule]{An illustrative example of a  relaxed CNF classification rule, which describes the decision function in the form of a  two-level checklist. This classifier is learned on the WDBC (Wisconsin diagnostic breast cancer) dataset and it predicts whether a tumor cell is malignant or not based on the characteristics of the  tumor cell. The first column in each checklist contains Boolean literals. An entry in the second column is  $ 1 $ if the corresponding literal is true by an observed tumor cell, and $ 0 $ otherwise. In the first level, checklist A (resp.\ B) is true if the count of true  literals is at least $ 2 $ (resp.\ $ 3 $). In the second level, a tumor cell is predicted as malignant if the count of true checklists is at least $ 1 $. }
	\label{interpretability_crr_exmpl:wdbc}
\end{figure*}

@@

{Through a comprehensive experimental evaluation over datasets from the UCI and Kaggle repository, we observe that {\crr} with relaxed-CNF rules achieves an improved trade-off between accuracy and rule sparsity and scales to datasets with more than $ 10^5 $ samples. More significantly, {\crr} generates relaxed-CNF rules with higher accuracy than CNF rules generated by~\cite{GM2019}. Furthermore, compared to decision lists generated by~\cite{cohen1995fast}, relaxed-CNF rules are sparser in large datasets.}
%can also reach the same level of accuracy compared to plain CNF/DNF rules with smaller relaxed-CNF rules. 
%
%We have conducted a comprehensive experimental study on the datasets from the UCI repository. Our empirical study demonstrates  an increase in the prediction accuracy for choosing relaxed-CNF rules over regular CNF rules. The incremental approach achieves runtime scalability without loss of accuracy.
